Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-09-05 redaction/slides_ETSG1.tex
Nouvell version des slides

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item Trade costs: A central role in international economic analysis \vspace{0.1cm}
\begin{itemize}
\item[-] Declining over the second half of the 20$^{th}$ century (Jacks et al., 2008, Novy, 2013) \vspace{0.1cm}
\item[-] But still significant: Average international trade costs = a 74\% markup over production costs (Anderson \& Van Wincoop, 2004)
\end{itemize}
\item What exactly are ``trade costs''?  \vspace{0.1cm}
\begin{itemize}
\item[-] Transaction costs, policy costs, time costs, and transport costs \textit{per se}
\end{itemize}
\item Transport costs: A sizeable share of international trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] Account for 21\% of international trade costs (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] If much trade policy barriers have been removed, the transport cost component of trade costs remains sizeable \vspace{0.1cm}
\end{itemize}
\textbf{The paper: On international transport costs}

@@

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Wouldn't trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
\item The structure (additive vs iceberg) of transport costs, far from being anecdotal \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Additive trade costs, an important role in shaping the pattern of trade flows (the Alchian \& Allen (1964) conjecture) \vspace{0.1cm}
\item[-] A bunch of empirical papers in support of the additive hypothesis: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
\item[-] With strong normative implications: Sorensen (2014), Irrarazabal et al. (2015) \vspace{0.1cm}}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.1cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}
\end{frame}

\begin{frame}
\frametitle{Our paper}
\begin{itemize}
\item An empirical decomposition of the structure of transport costs \vspace{0.1cm}
\item Quantitatively assess the size and the importance of the additive component in international transport costs
\begin{itemize}
%\footnotesize{
\item[-] Exploit the differences between export and imports prices (by transport mode, ocean and air)\vspace{0.1cm}
\item[-] From the US imports flows, on a yearly basis from 1974 to 2013  \vspace{0.1cm}
%}
\end{itemize}
\item Our contribution to the literature
\begin{itemize}
\item[-] Confirm the literature about the importance of the additive component of trade costs (Martin, 2012, Hummels \& Skiba, 2004)\vspace{0.1cm}
\item[-] But, we quantify it \vspace{0.1cm}
\item[-] Over a large spectrum of time (1974 to 2013) and distinguishing between air and sea transport \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] A broad view of the magnitude of additive costs in international trade over time
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Three questions, three answers}
\begin{itemize}
\item What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\item What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much \vspace{0.1cm}
\begin{itemize}
\item[-] With the additive term included, the ad-valorem component reduces by a factor of 2 \vspace{0.1cm}
\item[-] With the additive term taken into account, a significantly better ``goodness-of-fit'' \vspace{0.1cm}
\end{itemize}
\item How have international transport costs evolved over time? \vspace{0.1cm}
\begin{itemize}
\item[-] Transport costs \textit{per se} have fallen since 1985, by 40\% \vspace{0.1cm}
\item[-] When additive costs are included, not much difference between air and sea transport
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Plan of the talk}
\begin{itemize}
\item Data Sources  \vspace{0.1cm}
\item Empirical Methodology \vspace{0.1cm}
\item Assessing the importance of the additive component in transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Measure both the ad-valorem and the additive components  \vspace{0.1cm}
\item[-] Goodness-of-fit evaluation\vspace{0.1cm}
\end{itemize}
\item Characterizing the trends of transport costs \vspace{0.1cm}
\item Conclusion
\end{itemize}
\end{frame}

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
%\item Measure international transport costs as the difference between the export price and the import price \vspace{0.1cm}
\item Database: US annual imports of merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level \vspace{0.1cm}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the US import gate \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (1)}
\textbf{The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$ (and by transport mode, air or vessel) \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 (at the $k=$ 3-digit classification level)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification (2)}
\textbf{The estimation strategy}
\begin{itemize}
\item Two assumptions (as in Irrarazabal et al., 2015) \vspace{0.1cm}
\begin{itemize}
\item[-] Both iceberg and additive costs are separable between the origin country $i$ and the product $k$ dimensions \vspace{0.1cm}
\item[-] Separability in a multiplicative manner for ad-valorem costs and additive manner for per-kg costs \vspace{0.1cm}
\item[$\Leftrightarrow$] Write $t_{ik}$ and $\tau_{ik}$ as:
\begin{equation}
 \tau_{ik} = \tau_{i} \times \tau_{k}, \qquad t_{ik} = t_{i} + t_{k} \label{eq:specifTC}
 \end{equation}
\end{itemize}
\item Given the constraint $\frac{p_{ik}}{\widetilde{p}_{ik}} -1$, the error term should be always positive and multiplicative \vspace{0.1cm}
\item[$\Rightarrow$] The estimated equation becomes:
$$\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i} \times \tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})$$
\begin{itemize}
\item[-] With $\epsilon_{ik}$ following a normal law centered on 0.
\end{itemize}

@@

\begin{frame} [label=slide_method]
\begin{itemize}
\item Taking logs, we finally estimate the following equation
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:est_equation}
\end{equation}
\normalsize
\item A non-linear equation (due to the additive costs) $\rightarrow$ Estimation using non-linear squares \hyperlink{app_method}{\beamergotobutton{More}} \vspace{0.1cm}
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:est_equation}) constraining $t=0$ \vspace{0.1cm}
\begin{itemize}
\item[-] For each year and by transport mode, estimate two equations, depending on additive costs included or not \vspace{0.1cm}
\end{itemize}
\item When additive costs are included:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 +\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}\right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\normalsize
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{adv}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}, \qquad \widehat{t}^{add}_{ik} = \widehat{t}_{i} + \widehat{t}_{k}$$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item When assuming iceberg costs only
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\begin{itemize}
\item[-] From which we re-built:
$$\widehat{\tau}^{ice}_{ik} = \widehat{\tau_{i}} \times \widehat{\tau_{k}}$$
\end{itemize}
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{\tau}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Decomposing transport costs: Additive vs Iceberg}
\textbf{Result 1}: Provide estimates for the size of both the ad-valorem and the additive components of transport costs
\begin{itemize}
\item Transport costs expressed in percent of the export price, average values over 1974-2013 \hyperlink{app_results_summary}{\beamergotobutton{More}}  \vspace{0.1cm}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%

\begin{frame}
\frametitle{Comments}
\begin{itemize}
\item An important magnitude of overall transport costs  \vspace{0.1cm}
\begin{itemize}
\item[-] Over 1974-2013, correspond to a mean increase of the export price by 5.8\% for ocean shipping, by 5.1\% for air transport  \vspace{0.1cm}
\end{itemize}
\item When decomposing the structure of transport costs: A sizeable additive component   \vspace{0.1cm}
\begin{itemize}
\item[-] Additive costs amount to 2.9\% of the export price in ocean shipping, 1.8\% in air transport  \vspace{0.1cm}
\end{itemize}
\item Omitting the additive term substantially biases the iceberg component upwards  \vspace{0.1cm}
\begin{itemize}
\item[-] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[-] 3.2\% and 2.5\% as mean values in vessel and air respectively (in presence of additive costs) \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_fig1]
\frametitle{More on the importance of additive transport costs}
\begin{itemize}
\item Using the time dimension of our database, study the shares of both ad-valorem and additive components in total transport costs
\item[] \hyperlink{slide_moreonfig1}{\beamergotobutton{More on this figure later}}
\end{itemize}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
\includegraphics[width=5cm, height=5cm]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=5cm,height=5cm]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[label=slide_comments_fig1]
\frametitle{Comments}
\begin{itemize}
\item The size of transport costs is larger for ocean shipping than from air transport \vspace{0.1cm}
\begin{itemize}
\item[-] On average over the period (Table 1) and on a yearly basis (Figure 1) \hyperlink{app_comments_fig1}{\beamergotobutton{More}}  \vspace{0.1cm}
\end{itemize}
\item A sizeable share of the additive component in total transport costs  \vspace{0.1cm}
\begin{itemize}
\item[-] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[-] A decreasing share of additive costs over time for air transport \vspace{0.1cm}
\item[-] Roughly constant for ocean shipping
\end{itemize}
\end{itemize}
$\Rightarrow$ \textbf{Result 2}: Additive costs, an important dimension of international transport costs \vspace{0.1cm}
\begin{itemize}
\item Further investigate the robustness of this result
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{The role of additive costs: Goodness-of-fit evaluations}
\begin{itemize}
\item Provide a more systematic diagnosis about the importance of additive costs  \vspace{0.1cm}
\item By comparing the goodness-of-fit of the regressions
\begin{itemize}
\item[-] Obtained under Specification (\ref{eq:model_nlI}) (no additive costs)
\item[-] vs Specification (\ref{eq:model_with_add}) (with additive costs) \vspace{0.1cm}
\end{itemize}
\item Various measures of goodness of fit
\begin{itemize}
\item[-] The $R^2$ (the larger the value, the better the fit) \vspace{0.1cm}
\item[-] Standard Error of Regression (SER) (the smaller the value, the better the fit) \vspace{0.1cm}
\item[-] The log-likelihood function and two derived measures \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] The Akaike Information Criterion (the lower AIC, the better the fit) \vspace{0.1cm}
\item[$\ast$] The log-likelihood ratio test ($H_0$: both models are equivalent)
\end{itemize}
\end{itemize}

@@

\end{itemize}
 \hyperlink{slide_moreonfig1}{\beamergotobutton{Back to slide}}
\end{frame}

>>>>

\begin{frame}
\frametitle{Motivation}
\begin{itemize}
\item Trade costs: A central role in international economic analysis \vspace{0.1cm}
\begin{itemize}
\item[-] Declining over the second half of the 20$^{th}$ century (Jacks et al., 2008, Novy, 2013) \vspace{0.1cm}
\item[-] But still significant: Average international trade costs = a 74\% markup over production costs (Anderson \& Van Wincoop, 2004)
\end{itemize}
\item What exactly are ``trade costs''?  \vspace{0.1cm}
\begin{itemize}
\item[-] Transaction costs, policy costs, time costs, and transport costs \textit{per se}
\end{itemize}
\item Transport costs: A sizeable share of international trade costs \vspace{0.1cm}
\begin{itemize}
\item[-] Account for 21\% of international trade costs (Anderson \& Van Wincoop, 2004) \vspace{0.1cm}
%\item[-] Elasticity of trade wrt freight costs = -3 (Behar \& Venables, 2011) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] If much trade policy barriers have been removed, the transport cost component of trade costs remains sizeable \vspace{0.1cm}
\end{itemize}
\textbf{The paper: On international transport costs}

@@

\begin{frame}
\frametitle{Motivation (cont')}
\begin{itemize}
\item Standard modeling of trade costs: As an ad-valorem tax-equivalent \vspace{0.1cm}
\begin{itemize}
\item[-] As a constant percentage of the producer price per unit traded \vspace{0.1cm}
\item[$\Leftrightarrow$] The ``iceberg cost'' hypothesis (Samuelson, 1954) \vspace{0.1cm}
\end{itemize}
\item Yet... A debated question \vspace{0.1cm}
\item Wouldn't trade costs rather exhibit an additive structure ?  \vspace{0.1cm}
\item The structure (additive vs iceberg) of transport costs, far from being anecdotal \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Additive trade costs, an important role in shaping the pattern of trade flows: Martin (2012), Hummels \& Skiba (2004), Irrarazabal et al. (2015) \vspace{0.1cm}
\item[-] With strong normative implications: Sorensen (2014), Irrarazabal et al. (2015) \vspace{0.1cm}}
\end{itemize}
\item[$\Rightarrow$] Trade costs are likely to display an additive component, but precisely... by how much? \vspace{0.1cm}
\end{itemize}
\textbf{One objective of the paper: Provide an answer to this question}
\end{frame}

\begin{frame}
\frametitle{Our paper in 3 questions (and 3 answers)}
An empirical decomposition of the structure of transport costs \vspace{0.1cm}
\begin{itemize}
\item[(1)] What is the size of the iceberg and the additive trade costs? \vspace{0.1cm}
\item[$\Rightarrow$] Provide a quantitative measure of both  \vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air and ocean transport (mean value over 1974-2013) \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
}
\end{itemize}
\item[(2)] What do we lose by skipping the additive part of transport costs? \vspace{0.1cm}
\item[$\Rightarrow$] We lose much: With the additive term included, \vspace{0.1cm}
\begin{itemize}
\item[-] The estimated iceberg component reduces by a factor of 2 \vspace{0.1cm}
\item[-] A significantly better ``goodness-of-fit'' \vspace{0.1cm}
\end{itemize}
\item[(3)] How have international transport costs evolved over time? \vspace{0.1cm}
\begin{itemize}
\item[-] Transport costs \textit{per se} have fallen since 1985, by $\simeq$ 40\% \vspace{0.1cm}
\item[-] When additive costs are included, not much difference between air and sea transport, $\neq$ Hummels (2007) and Behar \& Venables (2011)
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Plan of the talk}
\begin{itemize}
\item Data Sources  \vspace{0.1cm}
\item Empirical Methodology \vspace{0.1cm}
\item Results \vspace{0.1cm}
\item Conclusion
\end{itemize}
\end{frame}

\begin{frame} [label=slide_data]
\frametitle{Data sources}
\begin{itemize}
\item Our measure of international transport costs: The difference between the export price and the import price \vspace{0.1cm}
\item Database: US Imports of Merchandise database \hyperlink{app_data}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] The export (fas) price, $\widetilde{p}$: the price for one kg of merchandise at the country export gate \vspace{0.1cm}
\item[-] The import (cif) price, $p$: the price for one kg of merchandise at the entry in the US \vspace{0.1cm}
\item[-] Yearly basis, from 1974 to 2013, HS 10 digit classification level, by transport mode (air or vessel) \vspace{0.1cm}
\end{itemize}
\item[$\Rightarrow$] Our dependent variable: The ratio $p/\widetilde{p}$ \vspace{0.1cm}
\begin{itemize}
\item[-] At the 3-digit classification level  \vspace{0.1cm}
\item[-] Estimation at the 4-digit level on some selected years as robustness \vspace{0.1cm}
\item[-] Approximatively 200 products (3 digits), from around 200 countries  \vspace{0.1cm}
\begin{itemize}
\item[$\ast$] Around 600-700 products at the 4-digit level
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Empirical specification: The estimated equation}
\begin{itemize}
\item Relate the import price $p$ to the export price $\widetilde{p}$ given both additive (per-kg) costs $t$ and ad-valorem costs $\tau$:
$$p = \tau \widetilde{p} + t, \qquad \text{with}\quad \tau \geq 1,\quad t \geq 0$$
\item For product $k$, from country $i$  \vspace{0.1cm}
\item Rewrite to get:
$$\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}}$$
\item[$\Rightarrow$] Estimate this equation for each year over 1974-2013 \vspace{0.1cm}
\begin{itemize}
\item[-] The equation being also year- and mode (air or vessel)- specific
\end{itemize}

@@

\begin{frame}  [label=slide_method]
\frametitle{Empirical specification: The estimation strategy }
\begin{itemize}
\item With some assumptions on the specification of transport costs \& the error term, and taking logs  \hyperlink{app_method_1}{\beamergotobutton{More}}\vspace{0.1cm}
\item[$\Rightarrow$] Estimate the following equation \vspace{0.1cm}
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_with_add}
\end{equation}
\begin{itemize}
\item[-] Estimates performed using non-linear least squares \hyperlink{app_method_2}{\beamergotobutton{More}} \vspace{0.1cm}
\end{itemize}
\normalsize
\item How to characterize the importance of additive costs relatively to iceberg? \vspace{0.1cm}
\item[$\Rightarrow$] Estimate Equation (\ref{eq:model_with_add}) constraining $t=0$:
\footnotesize
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \times \tau_{k}-1 \right) + \epsilon^{ice}_{ik} \label{eq:model_nlI}
\end{equation}
\normalsize
\item  Taking the average over the product-country dimension, we finally get (by year and transport mode): \hyperlink{app_method_2}{\beamergotobutton{More}}\vspace{0.1cm}
\begin{itemize}
\item[-] When additive costs are included: $\widehat{\tau}^{adv}$, $\widehat{t}^{add}$  \vspace{0.1cm}
\item[-] With only iceberg costs: $\widehat{\tau}^{ice}$
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components of transport costs (by year \& transport mode)\vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{ Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits ($^\ast$)} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Trade Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Trade Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
        \# obs. & 29279 & 28207 & 29317 & 27680 \\
   % \# origin country & 188 & 191 & 188 & 189 \\
   % \# products & 230 & 211 & 666 & 567 \\  \hline \hline
  \end{tabular}
    \end{center}}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{table}%
\begin{itemize}
\item Transport costs are sizeable \vspace{0.1cm}
\begin{itemize}
\item[-] Add $\simeq$ a 5\% margin over the export price \vspace{0.1cm}
\end{itemize}
\item Opening the black box of transport costs \vspace{0.1cm}
\begin{itemize}
\item[-] Iceberg cost: 2.5\% and 3.2\% of the export price in air \& ocean resp. \vspace{0.1cm}
\item[-] Additive cost: 1.8\% and 2.9\% of the export price  \vspace{0.1cm}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_result2]
\frametitle{Result 2: Additive transport costs do matter}
\begin{itemize}
\item International transport costs: A sizeable additive component \vspace{0.1cm}
\begin{itemize}
\item[-] Omitting the additive term substantially biases the iceberg component upwards (Table 1) \vspace{0.1cm}
\begin{itemize}
\item[$\star$] The ad-valorem cost reduces by a factor of 2 when additive transport costs are included in the estimation  \vspace{0.1cm}
\item[$\star$] Amounts to 3.2\% and 2.5\% as mean values in vessel and air respectively in presence of additive costs \vspace{0.1cm}
\item[$\star$] vs 5.8\% and 5.1\% when estimated alone \vspace{0.1cm}
\end{itemize}
\item[-] A sizeable share of the additive component in total transport costs \vspace{0.1cm}
\begin{itemize}
\item[$\star$] 48.2\% in average for ocean, 42.3\% for air  \vspace{0.1cm}
\item[$\star$] A result that holds throughout the period \hyperlink{slide_fig1}{\beamergotobutton{See Figure 1 later}} \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item A result confirmed by goodness of fit comparisons \vspace{0.1cm}
\begin{itemize}
\item[-] 4 measures of goodness of fit for Models (\ref{eq:model_with_add}) (with additive TC) and (\ref{eq:model_nlI}) (without additive TC) \vspace{0.1cm}
\item[$\Rightarrow$] A systematically better goodness of fit when including the additive component \hyperlink{app_goodnessfit}{\beamergotobutton{More results}}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label=slide_fig1]
\frametitle{Result 3: Characterizing the trends of transport costs }
\begin{itemize}
\item Study the shares of both ad-valorem and additive components in total transport costs  \hyperlink{slide_result2}{\beamergotobutton{Go back to Result 2}} \hyperlink{slide_comment_compositioneffects}{\beamergotobutton{Go back to Result 3}}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Air } & {\scriptsize  (b) Vessel}\\
\includegraphics[width=5cm, height=5cm]{Fig2a_decompTC_air_3d.pdf}
& \includegraphics[width=5cm,height=5cm]{Fig2b_decompTC_vessel_3d.pdf} \\
\end{tabular}
\end{center}
\end{figure}
\item[$\Rightarrow$] Downward trend in overall transport costs, for both air and ocean shipping since 1974 \hyperlink{app_fig1}{\beamergotobutton{Robustness of this result}}\vspace{0.1cm}
\begin{itemize}
\footnotesize{
\item[-] From 1980, a 1.5- 2 percentage points decrease in Air, a 2-3 percentage point decrease in Vessel }\vspace{0.1cm}
\end{itemize}
\
\end{itemize}
\end{frame}

\begin{frame}[label=slide_compositioneffects]
\frametitle{Time trends in transport costs: Excluding the composition effects }
\begin{itemize}
\item Does it mean a decrease in transport costs \textit{per se}? Not necessarily \vspace{0.2cm}
\item The change in overall transport costs over time: \vspace{0.1cm}
\begin{itemize}
\item[-] Depend on the evolution of per product- per partner costs, \vspace{0.1cm}
\item[-] But also on the composition of trade flows \vspace{0.2cm}
\begin{itemize}
\item[$\star$] Over time, import more goods that are cheaper to transport, and/or from countries with which it is cheaper to trade \vspace{0.1cm}
\end{itemize}
\end{itemize}
\item[$\Rightarrow$] Necessary to eliminate the composition effects of trade flows, to isolate the evolution of transport costs \textit{per se}   \vspace{0.1cm}
\item[$\Rightarrow$] What we do, in accordance with Hummels (2007) \hyperlink{app_compositioneffects}{\beamergotobutton{More on the method}} 
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Total transport costs (composition effects excluded) over time
\end{itemize}
\begin{figure}[htbp]
%\caption{Decomposing Transport costs (Yearly mean value, 3 digits)}
%\label{fig:decomp_TC_3d}
\begin{center}
\begin{tabular}{cc}
{\scriptsize (a) Iceberg alone } & {\scriptsize  (b) With additive }\\
\includegraphics[width=5cm, height=5cm]{Fig3a_TC_overtime_comp_effects_excl.pdf}
& \includegraphics[width=5cm,height=5cm]{TC_addplusmult_compeffects_excl.pdf} \\
\end{tabular}
\end{center}
\end{figure}

@@

\end{itemize}
 \hyperlink{slide_compositioneffects}{\beamergotobutton{Back to slide}}
\end{frame}
